---
title: "California Coastal Poll"
output:
  html_document:
    code_folding: hide
bibliography: ../coastal-perceptions.bib
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=F, warning = F)
```

## Overview

[@andersonPublicPerceptionsCoastal2012; @levinBiblicalInfluencesConservation2014]

## Read Data

- tidy principles
- cheatsheets, in RStudio Help or at  [Cheatsheets - RStudio](https://www.rstudio.com/resources/cheatsheets/):
    - [Data Import](https://github.com/rstudio/cheatsheets/raw/master/data-import.pdf)
    - [Data Transformation with dplyr](https://github.com/rstudio/cheatsheets/raw/master/data-transformation.pdf)
- paths: relative vs absolute
- data: raw (locked) vs derived, tmp

Want:

survey_id | heading | question | answer | value_num | value_chr
----------|---------|----------|--------|-----------|----------
...       |         |          |        |           |

```{r}
# libraries
library(tidyverse)
library(readxl)
library(here)
library(glue)

# paths
data_xlsx         <- here("data/raw/CoastalOpinionPoll_thru2017.xlsx")
headers_xlsx      <- here("data/derived/CoastalOpinionPoll_thru2017_headers.xlsx")
questions_csv     <- here("data/derived/questions.csv")
answers_csv       <- here("data/derived/answers.csv")
data_csv          <- here("data/derived/data.csv")
todo_chr2num_csv  <- here("data/derived/todo_data-not-numeric.csv")

headers  <- read_excel(headers_xlsx, col_types="text") %>% 
  gather(column, value, -row) %>% 
  spread(row, value) %>% 
  mutate(
    column = str_replace(column, fixed(".."), "") %>% as.numeric() - 1) %>% 
  arrange(column) %>% 
  select(column, heading, question, answer, comment1, comment2, comment3) %>% 
  fill(heading, question)
#View(headers)

headers %>% 
  group_by(heading, question) %>% 
  summarise(
    column1  = first(column),
    comment1 = first(comment1),
    comment2 = first(comment2),
    comment3 = first(comment3)) %>% 
  write_csv(questions_csv)
questions <- read_csv(questions_csv)

headers  <- headers %>% 
  select(-heading, -starts_with("comment"))
#View(headers)

n_max <- 12891-7
data <- read_excel(
  data_xlsx, 
  n_max=n_max, guess_max=n_max, skip=8, col_names=F)
#View(head(data))

col_class <- map_chr(data, class)
table(col_class)

data_chr <- data[, c(T, col_class[-1] %in% c("character"))]  %>% 
  rename(survey_id = "..1") %>% 
  gather(column, value_chr, -survey_id) %>% 
  mutate(
    column = str_replace(column, fixed(".."), "") %>% as.numeric()) %>%
  filter(!is.na(value_chr)) %>% 
  left_join(headers, by="column")
#View(data_chr)

# check to see if value_chr should be value_num
data_chr_ck <- data_chr %>% 
  group_by(column, question, answer) %>% 
  summarize(
    n = n())
# View(data_chr_ck)

# columns confirmed to be ok as character
cols_chr <- c(
  2,4:7,10:13,46,167,256,263,434,437,438,447,455,460,487)
# TODO: name NA answers: columns 256,438, 
# TODO: fix column 455 (QM) Zip Code : "answerd"
# - eg values of Leo Carrillo State Park row 10887, not 1 or 0

# convert character to numeric
data_chr_num <- data_chr %>% 
  filter(!column %in% cols_chr) %>% 
  mutate(
    value_num = as.numeric(value_chr))
# NAs introduced by coercion

# flag data to clean that didn't convert
data_chr_num %>% 
  filter(is.na(value_num)) %>% 
  write_csv(todo_chr2num_csv)

# remove converted numeric data from data_chr
data_chr <- data_chr %>% 
  filter(column %in% cols_chr)

# cleanup converted numeric data
data_chr_num <- data_chr_num %>% 
  select(-value_chr) %>% 
  filter(!is.na(value_num))

data_num <- data[, c(T, col_class[-1] %in% c("logical","numeric"))]  %>% 
  rename(survey_id = "..1") %>% 
  gather(column, value_num, -survey_id) %>% 
  mutate(
    column = str_replace(column, fixed(".."), "") %>% as.numeric()) %>%
  filter(!is.na(value_num)) %>% 
  left_join(headers, by="column")
#View(data_num)

d <- bind_rows(data_num, data_chr, data_chr_num) %>% 
  select(survey_id, column, question, answer, value_num, value_chr) %>%
  arrange(survey_id, column) %>% 
  filter(!str_detect(answer, "^x"))
#View(d)

# TODO: check for expected 1s or 0s

answers <- questions %>% 
  left_join(
    d, by="question") %>% 
  group_by(heading, question, answer, column) %>% 
  summarize(
    value_min = min(value_num, na.rm = T),
    value_max = max(value_num, na.rm = T),
    n_surveys = length(unique(survey_id))) %>% 
  ungroup() %>% 
  mutate(
    value_min = ifelse(is.infinite(value_min), NA, value_min),
    value_max = ifelse(is.infinite(value_max), NA, value_max)) %>% 
  arrange(column) %>% 
  write_csv(answers_csv)
answers <- read_csv(answers_csv)
#View(answers)
```


```{r old, eval=F}
d_qa <- questions %>% 
  select(heading, question) %>% 
  left_join(
    questions %>% 
      select(heading, question), 
    by="question") %>% 
  nest(survey_id, column, answer, value_num, value_chr)
#View(d)


d_n <- d %>% 
  left_join(
    questions %>% 
      select(heading, question), 
    by="question") %>% 
  nest(survey_id, column, answer, value_num, value_chr)
#View(d)

qs <- questions %>% 
  select(heading, question) %>% 
  nest(question) %>% 
  jsonlite::toJSON() %>%
  #listviewer::reactjson()
  listviewer::jsonedit()
```


## Coastal (General): CA ocean health better?

```{r}
library(lubridate)
here = here::here
library(RColorBrewer)
library(plotly)

q <- "CA ocean health better?"
o_q <- "Metadata"
o_a <- "year"
f_a <- "answered"

d_o <- d %>% 
  filter(
    question == !!o_q, 
    answer   == !!o_a) %>% 
  select(survey_id, value_num) %>% 
  rename(!!o_a := value_num)
d_o

d_q <- d %>% 
  filter(question == !!q) %>% 
  left_join(d_o, by="survey_id") %>% 
  arrange(survey_id, answer)


d_a <- d_q %>% 
  group_by(question, year, answer) %>% 
  summarize(
    sum = sum(value_num)) %>% 
  filter(answer != !!f_a) %>% 
  ungroup() %>% 
  mutate(
    year = ymd(year, truncated = 2))
d_a

# Stacked Percent
rdylgn <- brewer.pal(5,"RdYlGn")

g <- ggplot(d_a, aes(fill=answer, y=sum, x=year)) + 
  geom_bar( stat="identity", position="fill") +
  scale_fill_manual(values=c(rdylgn[1], "grey50", rdylgn[5])) +
  ylab("%") +
  theme_minimal()
g

ggplotly(g)
```


## Climate Change: Climate change problem?

Climate Change: perception over time, relationship with education and news?

```{r}
q    <- "Climate change problem?"
o_q  <- "Education"

pctbar_qyn_qm <- function(d, q, o_q, interactive=T){
  
  o_answers <- d %>% 
    filter(
      question == !!o_q, 
      answer   != "answered") %>% 
    select(column,answer) %>% 
    arrange(column) %>% 
    distinct() %>% 
    pull(answer)
  
  d_o <- d %>% 
    filter(
      question == !!o_q, 
      answer   != "answered",
      value_num == 1) %>% 
    select(survey_id, answer) %>% 
    rename(other = answer) %>% 
    mutate(
      other = factor(other, levels=o_answers, ordered=T))
  #d_o
  #d_o$other

  d_q <- d %>% 
    filter(
      question == !!q,
      answer   != "answered",
      value_num == 1) %>% 
    select(-value_num, -value_chr) %>% 
    left_join(d_o, by="survey_id") %>% 
    filter(!is.na(other)) %>% 
    arrange(survey_id, answer)
  #d_q
  
  d_a <- d_q %>% 
    group_by(question, answer, other) %>% 
    summarize(
      sum = n())
  #d_a
  
  # Stacked Percent
  rdylgn <- brewer.pal(5,"RdYlGn")
  
  g <- ggplot(d_a, aes(fill=answer, y=sum, x=other)) + 
    geom_bar( stat="identity", position="fill") +
    #facet_wrap(~condition) + 
    scale_fill_manual(values=c(rdylgn[1], "grey50", rdylgn[5])) +
    ylab("%") +
    xlab(o_q) +
    theme_minimal()
  
  if (interactive) return(ggplotly(g))
  g
}

pctbar_qyn_qm(d, "Climate change problem?", "Education")
```

```{r}
pctbar_qyn_qm(d, "Climate change problem?", "Env Issues")
```


## Recreational Activities: BBQ

### Treemap

[Draw Treemaps in 'ggplot2' • treemapify](http://wilkox.org/treemapify/)

```{r}
library(treemapify)

bbq <- d %>% 
  filter(
    question == "Recreational Activities",
    answer == "BBQ") %>% 
  pull(value_num)
#table(bbq)
# TODO: check value_num %in% c(0,1); ie fix value_num > 1

q <- "Recreational Activities"

d_a <- d %>% 
  filter(
    question == !!q,
    !answer %in% c("answered", "*response"),
    value_num > 0) %>% 
  group_by(answer) %>% 
  summarize(
    n = n())
#d_a

ggplot(d_a, aes(area = n, fill=answer, label=answer)) +
  geom_treemap() +
  geom_treemap_text(
    fontface = "italic", color = "white", 
    place = "centre", grow = TRUE) +   
  guides(fill=FALSE)
```

### Treemap animated by year

```{r}
library(gganimate)

q <- "Recreational Activities"

d_yr <- d %>% 
  filter(
    question == "Metadata", 
    answer   == "year") %>% 
  select(survey_id, year=value_num) %>% 
  mutate(
    year = as.integer(year))
#d_yr

d_a <- d %>% 
  filter(
    question == !!q,
    !answer %in% c("answered", "*response"),
    value_num > 0) %>% 
  left_join(
    d_yr, by = "survey_id") %>% 
  group_by(year, answer) %>% 
  summarize(
    n = n())
#d_a

g <- ggplot(d_a, aes(area = n, fill=answer, label=answer)) +
  geom_treemap() +
  geom_treemap(layout = "fixed") +
  geom_treemap_text(
    layout = "fixed", 
    fontface = "italic", color = "white", 
    place = "centre", grow = TRUE) +   
  guides(fill=FALSE) +
  transition_time(year) +
  ease_aes('linear') +
  labs(title = "Year: {frame_time}")

gif <- here(glue("figs/{q} animated_treemap.gif"))
anim_save(gif, g, nframes = 50)
```

![](`r gif`)


## References


